Manifold fitting by ridge estimation: a subspace-constrained approach

### 2021-02-25 Ridge_Estimation_main/Ridge-final.tex
zzh

\begin{abstract}
Modern statistics encounters with the high dimensional data in ambient space in almost everywhere. Although a data point usually represents itself as a long vector or a big matrix, in principle they all can be viewed as points on or near an intrinsic manifold. Estimation of the underlying manifold has been emerging as a great interest to the statistics community. In this paper, we propose the Subspace-Constrained Ridge Estimator (SCRE) to estimate the underlying manifold. The essence of the local SCRE is to extend a local version of the kernel density estimation (KDE) approach to the manifold fitting field. First, we demonstrate the connection between the ridge, essentially an approximation of the unknown manifold coming from the probability density function of the data, and the manifold itself.  Second, we show that, under mild conditions, this new approach has more promising theoretical results than the classical KDE approaches with respect to Hausdorff margin to the manifold. Third, we further show that the necessity of nonlinear transformation composed with the SCRE estimator, in order to improve the estimated ridge.  An algorithm is also provided which outputs an discrete ridge sets. Numerical simulated examples as well as real data sets demonstrate that the ridge obtained from our new approach does indeed have an improved average margin to the underlying manifold relative to other primary methods in the field of manifold fitting.
\end{abstract}

@@

Today, dealing with high-dimensional data is often costly, as it has required the exponential consumption of computational resources. Yet, even so, challenges remain, as the high-dimensional data tend to lie near a low dimensional sub-manifold of the ambient space, which is a phenomenon termed ``the curse of dimensionality'' by \cite{fefferman2018fitting}. Although, the geometric foundation of statistical inference in nonlinear regression was brought to attention of the statistical community as early as the 1980s by \cite{Kass1989}, the role of geometry was not primarily explored back then. At present, there are numerous works concerning manifold learning, with different areas of focus, such as dimension reduction focusing on approximating the embedding \cite{roweis2000nonlinear,zhang2004principal,donoho2003hessian} and manifold approximation via fitting the unknown manifold \cite{genovese2012manifold,chen2015asymptotic,genovese2014nonparametric}. In another seemingly-related direction, there have been works on statistical methods on manifolds over the past decades, centering on finding the mode of distribution \cite{huckemann2010intrinsic,chen2016nonparametric,chen2016comprehensive} and searching for the principal components of manifold \cite{hauberg2015principal,huckemann2006principal,hastie1989principal}.

There have also been some theoretical results regarding manifold learning using hypothesis testing. For instance, \cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. \cite{mohammed2017manifold} proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function ({\it asdf}) is orthogonal to the subspace spanned by the eigenvectors corresponding to the largest $D-d$ eigenvalues of the Hessian of {\it asdf}.

%Dimension reduction algorithms such as LLE, LTSA, Isomap\cite{roweis2000nonlinear,zhang2004principal,balasubramanian2002isomap} etc aim to find a new low-dimensional representation such that the local representation, local coordinate and geodesic distance are still retained to some degree within the lower dimensional space.

Manifold-fitting algorithms intend to recover the low-dimensional manifold in the ambient space by supposing that the observation data is generated from the manifold with some noise added. There are many classical algorithms to find the lower-dimensional manifold, such as the principal flow \cite{panaretos2014principal} and the local principal curves and surfaces based on kernel-density estimation \cite{ozertem2011locally}. The principal flow attempts to fit a curve moving along the maximal variation of the data, subject to a smoothness constraint. The solution to the principal flow is reduced to an ODE problem via the Euler-Lagrange method.  \cite{ozertem2011locally} redefined the principal curves and surfaces in terms of the gradient and Hessian of the probability density estimate. They also give an algorithm based on the classical KDE. \cite{jung2012analysis} propose a general framework to fit the high-dimensional data with principal nested spheres. In addition, there is a large amount of literature about searching for hidden structure with an application in the point clouds, nebular hypothesis, earthquake locations, and object-oriented objects \cite{adams2011morse,davenport2010joint,klemela2009smoothing,book}.

%{\color{red}Some works also address the bias reduction for the KDE method, for instance \cite{ImprovingBias2014,sain1996locally}. The exact behavior of the optimally adaptive smoothing parameter function is studied in \cite{sain1996locally}. The authors in \cite{ImprovingBias2014} reduce the bias by transforming a kernel into a higher-order kernel, via multiplication by polynomials. With the transformed kernel, they show that the bias can be made as small as possible and the variance be controlled not to increase.}

Overall, the problem of manifold estimation is treated as a statistical manifold-fitting problem under the observed data $\{x_i, i=1:n\}$. The observed data, as it may be assumed, are constructed as 
\[
x_i=\tilde{x}_i+\epsilon_i, i=1:n,
\] 
where noiseless data $\{\tilde{x}_i, i= 1:n\}$ are assumed to be sampled from some unknown manifold $\mathcal M$ and the noise $\epsilon_i$ obeys the rules of some distributions, such as the Gaussian distribution.  %With these observations, we can solve the manifold-interpolation problem to pull an outlier $x$ back to the manifold.  
Thus, the manifold-estimation problem is aimed at recovering a set of new samples $\{\hat{x}_i, i= 1:n'\}$ which we denoted as $\hat{\cal G}$ from the data $\{x_i\}$ such that the recovered samples in $\hat{\cal G}$ stay the real manifold $\mathcal M$ as much as possible. Normally, we can obtain $\hat{\cal G}$ by constructing an estimator using the data as
\begin{equation}\label{Estimator}
\hat{\cal G} = \hat{\cal G}(x_1,...,x_n | \Theta),
\end{equation}
where $\Theta$ represents the parameters used in the estimator. It is worth noting that the number of recovered samples in $\hat{\cal G}$ may be different from that of the data because we can always resample to find more points. Here, $\hat{\cal G}$ is named as a ridge estimator (discrete set). Regardless the methods/algorithms used to obtain the $\hat{\cal G}$, this estimator is, in principle, a subset of the ridge (continuous function) defined for the manifold $\cal{M}$ which, will elaborate the connection in the later sections.

The discussion on the relation of the principal curve, surface or ridge with the Hessian or gradient of the probability density function can be found in \cite{ozertem2011locally,genovese2014nonparametric,mohammed2017manifold}. In the low-dimensional structure estimation approach, KDE is commonly used in place of the unknown probability density function because of its concise form. The subspace-constrained-mean-shift algorithm is proposed to obtain the ridge numerically.

\subsection{The connection between $\cal M$ and the probability density function}
We use the probability density function as a useful tool to describe the manifold $\cal M$ for the following three reasons.
\begin{itemize}
\item The domain of the density function can be exactly on the manifold. 
\item The geometric property can be described by the analytic differential forms derived from the density function, such as gradient and Hessian matrix.
\item The probability of a point $x$ belongs to the manifold can be evaluated from the density function. As a result, we can recover the manifold from the density function by the sampling-rejection methods.
\end{itemize}
In this way, the manifold-fitting problem is closely related to the analysis of density function. Next, we will explain the reasons in details.
%In this paper, we provide a new algorithm to find the low-dimensional manifold using an iterative method. We also show that, in some cases, directly applying the iteration by projecting the gradient onto the D-d principal eigen-space of $H(x)$ is not a good choice because of the shifting eigenvalues of the matrix field $H(x)$. To overcome this difficulty, we select the proper eigenvectors from the eigen-space of $H(x)$ by evaluating each eigenvector with a function defined by the nearest tangent space from the current observation point $x$. Our method can be understood as a tangent-space-aided, subspace-constrained algorithm.

The manifold $\cal M$ can also be viewed as a particular lower-dimensional data distribution in the ambient space. On the population level, we can also use a density function $p(x)$ to describe the data distribution for $\cal M$ in the ambient space such as
%\[
%  p(x)= \left\{
%  \begin{array}{cc}
%  0, &x\not\in \cal M\\
%  \tilde{p}(x), &x\in \cal M
%  \end{array} 
%  \right. ,
%\]
\[
p(x)\geq 0,\quad  \int_{x\in \cal M} p(x)dx=1, \quad p(x) = 0, \forall x\not \in \cal M.
\] Noting that, even though $p(x)$ is not continuous for $x$ in the every direction (such as the directions in the normal space of $\cal M$), we can think of $p(x)$ as continuous and differentiable in the directions parallel with the manifold. Knowing $p(x)$ is equivalent for us to get the information of $\cal M$ because we can recover the manifold by re-sampling and rejecting the points having small value of $p(x)$.

Obviously, one can judge whether a point $x$ is on $\cal M$ by directly evaluating $p(x)$. The simplest way is to check the condition $p(x)\not = 0$, if true, $x\in \cal M$. Another useful characteristic is that, when $x\in \cal M$, we know the gradient at $x$ exactly lies in the tangent space of $\cal M$ at $x$, i.e., $\nabla p(x)\in {\cal T}_x$. For the density function $p(x)$, we can get the tangent space from the eigenvalue decomposition of the subspace-constrained Hessian matrix of $p(x)$ at $x$. The discontinuous property of $p(x)$ makes it impossible to derive the derivative in the normal directions of $\cal M$. To overcome this problem, we use the smooth density function $p_h(x)$ through convolution as follows
\begin{equation}\label{smoothed density}
p_h(x) = \int p(y) K_h(x-y)dy.
\end{equation}
Noting the resultant density function $p_h(x)$ is continuous and smooth in every direction. We would also need to choose a proper $h$ such that the following conditions satisfying:
\begin{itemize}
\item For $x\not \in \cal M$, the gradient $\nabla p_h(x)$ points to ${P_{\cal M}}(x)$ (the projected nearest point of $x$ in $\cal M$).
\item For $x\in \cal M$, the gradient $\nabla p_h(x)$ lies in the space spanned by the top $d$ eigenvectors of $H_{p_h}(x)$, i.e., the tangent space of $\cal M$ at $x$.
\item There is $\tau$, such that when $x\in \cal M\oplus \tau$, the projection $P(x) = U(x)U(x)^T$ corresponding to the top $d$ eigenvectors of the Hessian $p_h(x)$ change continuously when $x\rightarrow {P_{\cal M}}(x)$, where ${\cal M}\oplus \tau = \cup_{y\in \cal M} B^D_\tau (y)= \cup_{y\in \cal M} \{x|\|x-y\|_2\leq \tau\}$.
\end{itemize}

On the sample level, we will need to use the empirical version instead of \eqref{smoothed density} as
\[
 \hat{p}_{n,h}(x) = \frac{1}{n}\sum_{k=1}^n K_h(x-x_k).
\]
where $h$ is the bandwidth and $n$ is the number of samples. In order to ensure $\int  \hat{p}_{n,h}(x) dx = 1$, we need to normalize it within $K_h(x-x_k)$. Usually, when we do not distinguish the bandwidth in each dimension, $K_h(x-x_k)$ can yield a simple KDE form as ${1}/{h^D} K({x_i-x}/{h})$. Assume $K(u)$ is a kernel which satisfies
\[
\int K(u) du = 1, \int u K(u) du = 0, \int uu^T K(u) du= I \{{1}/{D} \int \|u\|_2^2 K(u) du\} .
\]
For ease of notation, from now on, we will omit the subscript $n$ in $ \hat{p}_{n,h}(x)$ as $\hat{p}_h(x)$, when we do not need to emphasize $n$.

\begin{remark}   
 $\hat{p}_h(x)$ is a random function depends on the observations $\{x_k\}$. Using $\hat{p}_h(x)$ to estimate $p_h(x)$ is unbiased, i.e., the expectation of $\hat{p}_h(x)$ is $p_h(x)$. Even though we can use $ \hat{p}(x)$ to approximate $p_h(x)$, it is not guaranteed that the gradient and Hessian of $\hat{p}_h(x)$ will approximates those of $p(x)$ well enough, such as shown in the work \cite{sasaki2017mode}.
  It is well-known that ${\rm E}(\hat{p}_h(x))-p(x) = O(h^2)$, which is also true for the derivatives.
\end{remark}

\subsection{Subspace-Constrained Derivatives}
Note that $p(x)$ is only smooth and differentiable in the direction paralleled with $\cal M$. For $v\in {\cal T}_x$, recall that the directional derivative of $p(x)$ is
\[
 \partial_v p(x) = \lim_{t\rightarrow 0} \frac{p(x+vt)-p(x)}{t}.
\]
Similarly, we have the second order directional derivative $\partial_{v_1,v_2}p(x)$ by recurrently differentiate from the first order derivative. For the tangent space ${\cal T}_x$ of $\cal M$ at $x$, we can find any orthogonal basis $V(x) = \{v_1(x), v_2(x),..., v_d(x)\}$. Then the Hessian constrained in the tangent space is
\[
H_{ V}(x) = V(x) X V(x)^T.
\]
where $X$ is a $d\times d$ matrix with the $i,j$-th element as $X_{ij}=  \{ \partial_{v_i,v_j} p(x )\}, 0\leq i,j \leq d$. Since $X$ is a real symmetric matrix, we can find the eigenvalue decomposition 
\[
X = U(x) \Lambda U(x)^T.
\]
where $U(x)$ is an $d\times d$ orthonormal matrix and $\Lambda$ is a $d\times d$ diagonal matrix. Next, when we use the columns of $M(x) = V(x)U(x)$ as the new basis in the tangent space, we will get 
\[
H_{V}(x) = M(x) \Lambda M(x)^T.
\]
After defining the subspace constrained gradient and Hessian, we can extent the non-zero domain of $p(x)$ to  $\cal M\oplus \tau$ such as
\[
\bar{p}(x) = p(\tilde{x}) + (x-\tilde{x})^T M(\tilde{x}) M^T(\tilde{x}) \nabla p_M (\tilde{x}) +\frac{1}{2}  (x-\tilde{x})^TM(\tilde{x})\Lambda M(\tilde{x})^T (x-\tilde{x}).
\]
The constructed $\bar{p}(x)$ has several good properties: 
\begin{itemize}
\item Firstly, for any $x_1,x_2 \in \cal M\oplus \tau$, if $P_{\cal M}(x_1) = P_{\cal M}(x_2)$, we have $\bar{p}(x_1) = \bar{p}(x_2)$. 
\item Secondly, the gradient and Hessian keep unchanged with $x$ moving in the normal space which is a constant vector or matrix equaling to the subspace-constraint gradient and Hessian of $p(x)$ at the projected point $P_{\cal M}(x)$.
\end{itemize}

To be more concrete, the set corresponding to large $\bar{p}(x)$ will produce a geometric structure $\bar{\cal M}$ much thicker than $\cal M$. Even though we can extend the non-zero domain of $p(x)$ to get $\bar{p}(x)$, $\bar{p}(x)$ is still not continuous and smooth on the boundary of $\cal M\oplus \tau$.

\subsection{Hessian Matrix Decomposition for $p_h(x)$}
For this reason, we can consider the convolution form $p_h(x)$ which is smooth and continuous everywhere.

It is easily to see that $p_h(x) = p(x)+O(h^2)$ by a few steps of calculating.
% By taking derivatives on both sides, we have, after the convolutional operation, the Hessian matrix of $p_h(x)$ becomes
%\[
% H_{p_h}(x) = M \Lambda M^T + O(h^2)
%\]
Because $p(x)$ only has derivatives along the manifold when $x\in \cal M$ and $p_h(x)$ is differentiable in every direction, we need to split the Hessian of $p_h(x)$ into two parts based on the tangent space of $p(x)$. That is, we consider the derivatives in two independent spaces respectively. For the Hessian in the tangent space, we have
\[
\begin{aligned}
 &M(x)^T H_{p_h}(x) M(x) =  \Lambda + O(h^2) \mathbbm 1_{d\times d},\\
 &M(x)_\perp^T H_{p_h}(x) M(x)_\perp  =  O(h^2) {\mathbbm 1}_{\{D-d\}\times \{D-d\}},\\
 &M(x)_\perp^T H_{p_h}(x) M(x) = O(h^2) {\mathbbm 1}_{\{D-d\}\times d}.
\end{aligned}
\]
Merging the above three equalities, we conclude that $H_{p_h}(x)$ can be written as a main part and a small part concerning $h$ by
\[
H_{p_h}(x) = M(x) \Lambda M(x)^T + O(h^2){\mathbbm 1}_{D\times D}.
\]
The existing of the perturbation term $O(h^2){\mathbbm 1}_{D\times D}$ in $H_{p_h}(x)$ makes it difficult to recover the tangent space orthonormal matrix $M(x)$ directly when $h$ turning large gradually. To fix this problem, in this paper, we propose a local approach to construct a better estimator and explain the effectiveness of our approach in reducing the effect of the perturbation. Our approach recovers the subspace $M(x)$ more accurately.

\subsection{Ridge and Manifold}
The most common examples of the manifold $\cal M$ are $d$-dimensional structures embedded in a $D$-dimensional space, such as $\cal M$ being the ball surface ($d=2, D=3$) or the torus ($d=2, D=3$). To handle the manifold situation, we define a function 
\[\phi(x): {\mathbb R}^D\rightarrow {\mathbb R}^+,
\] 
and construct a set 
\[R = \{x | \omega (\phi(x),\nabla \phi(x), H_\phi(x))=0 \}\] 
such that $R$ will approximate $\cal M$ well. Here, we refer $\omega$ to some abstract relations or formula. When $\phi(x)$ is selected as the density function, then, the points on $\cal M$ satisfy the subspace-constrained optimal condition. Because, for $x\in \cal M$, $\phi(x)$ will decrease when $x$ moves along any direction $v$ as long as $v\perp {\cal T}_{\cal M}(x)$, i.e, $x$ is a local optimal for the normal space of $\cal M$ at $x$. We call the set $R$ constituting of points satisfying the subspace-constrained optimal condition the {\it Ridge}, which is similar as the ridge of a mountain in our daily life.

%Furthermore, by using the more accurately estimated tangent subspace, we could obtain a ridge more closer with the real manifold.

@@

The distances, no matter Hausdorff or average margin, are complicated to bound or optimize directly. Because $\hat{\cal G}\subset \hat{R}$,  we need also connect $\cal M$ with some ridge by the following assumption:

@@

Specifically, the  distance between $R_1$ and $R_2$ is the same order with the error of the corresponding derivatives. 

>>>>

\begin{abstract}
Modern statistics encounters with the high dimensional data in ambient space in almost everywhere. Although a data point usually represents itself as a long vector or a big matrix, in principle they all can be viewed as points on or near an intrinsic manifold. Estimation of the underlying manifold has been emerging as a great interest to the statistics community. In this paper, we propose the Subspace-Constrained Ridge Estimator (SCRE) to estimate the underlying manifold. The essence of the local SCRE is to extend a local version of the kernel-density estimation (KDE) approach to the manifold fitting field. First, we demonstrate the connection between the ridge, essentially an approximation of the unknown manifold coming from the probability density function of the data, and the manifold itself.  Second, we show that, under mild conditions, this new approach has more promising theoretical results than the classical KDE-based approaches with respect to Hausdorff margin to the manifold. Third, we further show that the necessity of nonlinear transformation composed with the SCRE estimator, in order to improve the estimated ridge.  An algorithm is also provided which outputs an discrete ridge sets. Numerical simulated examples as well as real data sets demonstrate that the ridge obtained from our new approach does indeed have an improved average margin to the underlying manifold relative to other primary methods in the field of manifold fitting.
\end{abstract}

@@

Today, the process of dealing with high-dimensional data requires the exponential consumption of computational resources, and is therefore costly but effective. Yet, even so, challenges remain, as the high-dimensional data tend to lie near a low-dimensional sub-manifold of the ambient space, which is a phenomenon termed ``the curse of dimensionality'' by \cite{fefferman2018fitting}.  Although, the geometric foundation of statistical inference in nonlinear regression was brought to attention of the statistical community as early as the 1980s by \cite{Kass1989}, the role of geometry was not primarily explored back then. At present, there are numerous works concerning manifold learning, with different areas of focus, such as dimension reduction focusing on approximating the embedding \citep{roweis2000nonlinear,zhang2004principal,donoho2003hessian,zha2007} and manifold approximation via fitting the unknown manifold \citep{genovese2012manifold,chen2015asymptotic,genovese2014nonparametric}. In another seemingly-related direction, there have been works on statistical methods on manifolds over the past decades, centering on finding the mode of distribution \citep{huckemann2010intrinsic,chen2016nonparametric,chen2016comprehensive} and searching for the principal components of manifold \citep{hauberg2015principal,huckemann2006principal,hastie1989principal}.

There has also been some theoretical research into manifold learning using hypothesis testing. For instance, \cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. Mohammed and Narayanan proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function (asdf) is orthogonal to the subspace spanned by the largest $D-d$ eigenvectors of the Hessian of the asdf \citep{mohammed2017manifold}.

%Dimension-reduction algorithms such as LLE, LTSA, Isomap\cite{roweis2000nonlinear,zhang2004principal,balasubramanian2002isomap} etc. aim to find a new low-dimensional representation such that the local representation, local coordinate, and geodesic distance are still retained to some degree within the lower-dimensional space.

Manifold-fitting algorithms intend to recover the low-dimensional manifold in the ambient space by supposing that the observation data is generated from the manifold with some noise added. There are many classical algorithms to find the lower-dimensional manifold, such as the principal flow \citep{panaretos2014principal} and the local principal curves and surfaces based on kernel-density estimation \citep{ozertem2011locally}. The principal flow attempts to fit a curve moving along the maximal variation of the data, subject to a smoothness constraint. The solution to the principal flow is reduced to an ODE problem via the Euler-Lagrange method.  \cite{ozertem2011locally} redefined the principal curves and surfaces in terms of the gradient and Hessian of the probability density estimate. They also provided an algorithm based on the classical KDE. \cite{jung2012analysis} propose a general framework to fit the high-dimensional data with principal nested spheres. In addition, there is a large amount of literature about searching for hidden structure with an application in the point clouds, nebular hypothesis, earthquake locations, and object-oriented objects \citep{adams2011morse,davenport2010joint,klemela2009smoothing,book}.

%Some works also address the bias reduction for the KDE method, for instance \cite{ImprovingBias2014,sain1996locally}. The exact behavior of the optimally adaptive smoothing parameter function is studied in \cite{sain1996locally}. \cite{ImprovingBias2014} reduced the bias by transforming a kernel into a higher-order kernel, via multiplication by polynomials. With the transformed kernel, they showed that the bias could be made as small as possible and the variance be controlled to not increase.

Overall, the problem of manifold estimation is treated as a statistical manifold-fitting problem under the observed data $\{x_i, i=1:n\}$. The observed data, it may be assumed, are constructed as 
\[
x_i=\tilde{x}_i+\epsilon_i, i=1:n,
\] 
where noiseless data $\{\tilde{x}_i, i= 1:n\}$ are assumed to be sampled from some unknown manifold $\mathcal M$ and the noise $\epsilon_i$ obeys the rules of some distributions, such as the Gaussian distribution.  %With these observations, we can solve the manifold-interpolation problem to pull an outlier $x$ back to the manifold.  
Thus, the manifold-estimation problem is aimed at recovering a set of new samples $\{\hat{x}_i, i= 1:n'\}$, which we label $\hat{\cal G}$ from the data $\{x_i\}$ such that the recovered samples in $\hat{\cal G}$ stay in the real manifold $\mathcal M$ as much as possible. Normally, we can obtain $\hat{\cal G}$ by constructing an estimator using the data:
\begin{equation}\label{Estimator}
\hat{\cal G} = \hat{\cal G}(x_1,...,x_n | \Theta),
\end{equation}
where $\Theta$ represents the parameters used in the estimator. It is worth noting that the number of recovered samples in $\hat{\cal G}$ may be different from that of the data because we can always resample to find more points. Here, $\hat{\cal G}$ is named as a ridge estimator (discrete set). Regardless of the methods/algorithms used to obtain $\hat{\cal G}$, this estimator is, in principle, a subset of the ridge (continuous function) defined for the manifold $\cal{M}$; the connections between them will be elaborated in the later sections.

\subsection{The connection between $\cal M$ and the probability density function}
We use the probability density function as a useful tool to describe the manifold $\cal M$ for the following three reasons.
\begin{itemize}
\item The domain of the density function can be exactly on the manifold. 
\item The geometric property can be described by the analytic differential forms derived from the density function, such as the gradient and Hessian matrix.
\item The probability that a point $x$ belongs to the manifold can be determined from the density function. As a result, we can recover the manifold from the density function through sampling-rejection methods.
\end{itemize}

\begin{remark}   
The discussion on the relation among the principal curve/surface, the ridge with the Hessian or the gradient of the probability density function can be found in \cite{ozertem2011locally,genovese2014nonparametric,mohammed2017manifold}. In the low-dimensional structure estimation approach, KDE, as a plug-in method, is commonly used in place of the unknown probability density function because of its concise form. %The subspace-constrained-mean-shift algorithm is proposed to obtain the ridge numerically.
\end{remark}

In this way, the manifold-fitting problem is closely related to the analysis of the density function. We will now explain the reasons in details.
%In this paper, we provide a new algorithm to find the low-dimensional manifold using an iterative method. We also show that, in some cases, directly applying the iteration by projecting the gradient onto the D-d principal eigen-space of $H(x)$ is not a good choice because of the shifting eigenvalues of the matrix field $H(x)$. To overcome this difficulty, we select the proper eigenvectors from the eigen-space of $H(x)$ by evaluating each eigenvector with a function defined by the nearest tangent space from the current observation point $x$. Our method can be understood as a tangent-space-aided, subspace-constrained algorithm.

The manifold $\cal M$ can also be viewed as a particular lower-dimensional data distribution in the ambient space. At the population level, we can also use a density function $p(x)$ to describe the data distribution for $\cal M$ in the ambient space:
%\[
%  p(x)= \left\{
%  \begin{array}{cc}
%  0, &x\not\in \cal M\\
%  \tilde{p}(x), &x\in \cal M
%  \end{array} 
%  \right. ,
%\]
\[
p(x)\geq 0,\quad  \int_{x\in \cal M} p(x)dx=1, \quad p(x) = 0, \forall x\not \in \cal M.
\] 
It should be noted that, even though $p(x)$ is not continuous for $x$ in every direction (such as the directions in the normal space of $\cal M$), we can think of $p(x)$ as continuous and differentiable in the directions parallel with the manifold. Knowing $p(x)$ is equivalent to obtaining the information about $\cal M$, because we can recover the manifold by re-sampling and rejecting the points with small values of $p(x)$.

Clearly, it is possible to judge whether a point $x$ is on $\cal M$ by directly evaluating $p(x)$. The simplest way is to check the condition $p(x)\not = 0$: if true, $x\in \cal M$. Another useful characteristic is that, when $x\in \cal M$, we know the gradient at $x$ lies exactly in the tangent space of $\cal M$ at $x$, i.e. $\nabla p(x)\in {\cal T}_x$. For the density function $p(x)$, we can get the tangent space from the eigenvalue decomposition of the subspace-constrained Hessian matrix of $p(x)$ at $x$. The discontinuous property of $p(x)$ makes it impossible to derive the derivative in the normal directions of $\cal M$. To overcome this problem, we use the smooth density function $p_h(x)$ through convolution:
\begin{equation}\label{smoothed density}
p_h(x) = \int p(y) K_h(x-y)dy.
\end{equation}
Note that the resultant density function $p_h(x)$ is continuous and smooth in every direction. We would also need to choose a proper $h$ such that the following conditions are satisfied:
\begin{itemize}
\item For $x\not \in \cal M$, the gradient $\nabla p_h(x)$ points to ${P_{\cal M}}(x)$ (the projected nearest point of $x$ in $\cal M$).
\item For $x\in \cal M$, the gradient $\nabla p_h(x)$ lies in the space spanned by the top $d$ eigenvectors of $H_{p_h}(x)$, i.e., the tangent space of $\cal M$ at $x$.
\item There is $\tau$, such that when $x\in \cal M\oplus \tau$, the projection $P(x) = U(x)U(x)^T$ corresponding to the top $d$ eigenvectors of the Hessian $p_h(x)$ changes continuously when $x\rightarrow {P_{\cal M}}(x)$, where ${\cal M}\oplus \tau = \cup_{y\in \cal M} B^D_\tau (y)= \cup_{y\in \cal M} \{x|\|x-y\|_2\leq \tau\}$.
\end{itemize}

At the sample level, we will need to use the empirical version, instead of \eqref{smoothed density}, as follows:
\[
 \hat{p}_{n,h}(x) = \frac{1}{n}\sum_{k=1}^n K_h(x-x_k).
\]
where $h$ is the bandwidth and $n$ is the number of samples. In order to ensure $\int  \hat{p}_{n,h}(x) dx = 1$, we need to normalize it within $K_h(x-x_k)$. Usually, when we do not distinguish the bandwidth in each dimension, $K_h(x-x_k)$ can yield a simple KDE form, ${1}/{h^D} K({x_i-x}/{h})$. Assume $K(u)$ is a kernel that satisfies
\[
\int K(u) du = 1, \int u K(u) du = 0, \int uu^T K(u) du= I \{{1}/{D} \int \|u\|_2^2 K(u) du\} .
\]
For ease of notation, from this point we will omit the subscript $n$ in $ \hat{p}_{n,h}(x)$ (i.e. render it as $\hat{p}_h(x)$), when we do not need to emphasize $n$.

\begin{remark}   
 $\hat{p}_h(x)$ is a random function depending on the observations $\{x_k\}$. $\hat{p}_h(x)$ can be used to estimate $p_h(x)$ is unbiased, i.e. the expectation of $\hat{p}_h(x)$ is $p_h(x)$. Even though we can use $ \hat{p}(x)$ to approximate $p_h(x)$, it is not guaranteed that the gradient and Hessian of $\hat{p}_h(x)$ will approximate those of $p(x)$ well enough, as shown in the work of \cite{sasaki2017mode}.
  It is well known that ${\rm E}(\hat{p}_h(x))-p(x) = O(h^2)$, and this is also true for the derivatives.
\end{remark}

\subsection{Subspace-Constrained Derivatives}
Note that $p(x)$ is only smooth and differentiable in the direction parallel with $\cal M$. For $v\in {\cal T}_x$, recall that the directional derivative of $p(x)$ is
\[
 \partial_v p(x) = \lim_{t\rightarrow 0} \frac{p(x+vt)-p(x)}{t}.
\]
Similarly, we have the second-order directional derivative $\partial_{v_1,v_2}p(x)$ by recurrently differentiating from the first-order derivative. For the tangent space ${\cal T}_x$ of $\cal M$ at $x$, we can find any orthogonal basis $V(x) = \{v_1(x), v_2(x),..., v_d(x)\}$. Then the Hessian constrained in the tangent space is
\[
H_{ V}(x) = V(x) X V(x)^T.
\]
where $X$ is a $d\times d$ matrix with the $i,j$-th element as $X_{ij}=  \{ \partial_{v_i,v_j} p(x )\}, 0\leq i,j \leq d$. Since $X$ is a real symmetric matrix, we can find the eigenvalue decomposition 
\[
X = U(x) \Lambda U(x)^T.
\]
where $U(x)$ is an $d\times d$ orthonormal matrix and $\Lambda$ is a $d\times d$ diagonal matrix. Next, when we use the columns of $M(x) = V(x)U(x)$ as the new basis in the tangent space, we will get 
\[
H_{V}(x) = M(x) \Lambda M(x)^T.
\]
After defining the subspace-constrained gradient and Hessian, we can extent the non-zero domain of $p(x)$ to  $\cal M\oplus \tau$:
\[
\bar{p}(x) = p(\tilde{x}) + (x-\tilde{x})^T M(\tilde{x}) M^T(\tilde{x}) \nabla p_M (\tilde{x}) +\frac{1}{2}  (x-\tilde{x})^TM(\tilde{x})\Lambda M(\tilde{x})^T (x-\tilde{x}).
\]
The constructed $\bar{p}(x)$ has several good properties: 
\begin{itemize}
\item First, for any $x_1,x_2 \in \cal M\oplus \tau$, if $P_{\cal M}(x_1) = P_{\cal M}(x_2)$, we have $\bar{p}(x_1) = \bar{p}(x_2)$. 
\item Second, the gradient and Hessian remain unchanged with $x$ moving in the normal space, which is a constant vector or matrix equal to the subspace-constrained gradient and Hessian of $p(x)$ at the projected point $P_{\cal M}(x)$.
\end{itemize}

To be more concrete, the set corresponding to a large $\bar{p}(x)$ will produce a geometric structure $\bar{\cal M}$ much thicker than $\cal M$. Even though we can extend the non-zero domain of $p(x)$ to get $\bar{p}(x)$, $\bar{p}(x)$ is still not continuous and not smooth on the boundary of $\cal M\oplus \tau$.

\subsection{Hessian Matrix Decomposition for $p_h(x)$}
For this reason, we can consider the convolution form $p_h(x)$, which is smooth and continuous everywhere.

A few calculation steps make it clear that $p_h(x) = p(x)+O(h^2)$.
% By taking derivatives on both sides, we have, after the convolutional operation, the Hessian matrix of $p_h(x)$ becoming
%\[
% H_{p_h}(x) = M \Lambda M^T + O(h^2)
%\]
Because $p(x)$ only has derivatives along the manifold when $x\in \cal M$ and $p_h(x)$ are differentiable in every direction, we need to split the Hessian of $p_h(x)$ into two parts based on the tangent space of $p(x)$. In other words, we consider the derivatives in two independent spaces respectively. For the Hessian in the tangent space, we have
\[
\begin{aligned}
 &M(x)^T H_{p_h}(x) M(x) =  \Lambda + O(h^2) \mathbbm 1_{d\times d},\\
 &M(x)_\perp^T H_{p_h}(x) M(x)_\perp  =  O(h^2) {\mathbbm 1}_{\{D-d\}\times \{D-d\}},\\
 &M(x)_\perp^T H_{p_h}(x) M(x) = O(h^2) {\mathbbm 1}_{\{D-d\}\times d}.
\end{aligned}
\]
Merging the above three equalities, we conclude that $H_{p_h}(x)$ can be written as a main part and a small part concerning $h$:
\[
H_{p_h}(x) = M(x) \Lambda M(x)^T + O(h^2){\mathbbm 1}_{D\times D}.
\]
The existence of the perturbation term $O(h^2){\mathbbm 1}_{D\times D}$ in $H_{p_h}(x)$ makes it difficult to recover the tangent-space orthonormal matrix $M(x)$ directly when $h$ gradually becomes large. To fix this problem, in this paper we propose a local approach to construct a better estimator and explain the effectiveness of our approach in reducing the effect of the perturbation. Our approach recovers the subspace $M(x)$ more accurately.

\subsection{Ridge and Manifold}
The most common examples of the manifold $\cal M$ are $d$-dimensional structures embedded in a $D$-dimensional space, such as $\cal M$ being the ball surface ($d=2, D=3$) or the torus ($d=2, D=3$). To deal with the manifold, we define a function 
\[\phi(x): {\mathbb R}^D\rightarrow {\mathbb R}^+,
\] 
and construct a set 
\[R = \{x | \omega (\phi(x),\nabla \phi(x), H_\phi(x))=0 \}\] 
such that $R$ will approximate $\cal M$ well. Here, we refer $\omega$ to some abstract relations or formulas. When $\phi(x)$ is selected as the density function, the points on $\cal M$ satisfy the subspace-constrained optimal condition. This is because, for $x\in \cal M$, $\phi(x)$ will decrease when $x$ moves along any direction $v$ as long as $v\perp {\cal T}_{\cal M}(x)$, i.e. $x$ is a local optimal for the normal space of $\cal M$ at $x$. We call the set $R$ consisting of points satisfying the subspace-constrained optimal condition the {\it Ridge}, which is similar to the ridge of a mountain in the real world.

%Furthermore, by using the more accurately estimated tangent subspace, we can obtain a ridge closer to the real manifold.

@@

The distances, whether Hausdorff or average margin, are too complicated to bound or optimize directly. Because $\hat{\cal G}\subset \hat{R}$,  we also need to connect $\cal M$ with some ridge under the following assumption:

@@

Specifically, the  distance between $R_1$ and $R_2$ is of the same order as the error of the corresponding derivatives. 
